---
title: "R key list of useful commands"
author: "Saulai Vue"
date: "2025-04-03"
output: word_document
---

```{r setup, echo=FALSE, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This list is mainly used to combine all operators, functions, and useful commands in R primarily for practicing in class. Note some straightforward elements I have not put here.

## R Arithmetic Operators
%%  Modulus (Remainder from Division)
%/%  Integer Division

## Assignment Operators
"my_var <- 3"  "my_var <<- 3"

## R Comparison Operators
==  Equal
!=  Not equal
>   Greater than
<   Less than
>=  Greater than or equal to
<=  Less than or equal to

## R Logical Operators
&   Element-wise Logical AND operator. Returns TRUE if both elements are TRUE.
&&  Logical AND operator = Returns TRUE if both statements are TRUE.
|   Element-wise Logical OR operator. Returns TRUE if one of the statements is TRUE.
||  Logical OR operator. Returns TRUE if one of the statements is TRUE.
!   Logical NOT - Returns FALSE if statement is TRUE

## R Miscellaneous Operators
:   Creates a series of numbers in a sequence   "x <- 1:10"
%in%  Find out if an element belongs to a vector   "x %in% y" --- Returns a logical vector

## List of Package Libraries
```{r , echo=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(tidyverse)
library(ggplot2)
library(ggrepel)
library(gganimate)
library(tidyverse)
library(gapminder) #for some data
library(gifski) #creation of gifs, could also use library(png)
library(lubridate) #working with time variables
library(maps) #for mapping with ggplot
library(mapdata) #for mapping with ggplot
library(transformr) #for changing chloropleths across time
library(ggthemes) #extra color options

```

## Github
```{r , echo=FALSE, warning=FALSE}
usethis::git_sitrep()
# https://gist.github.com/Z3tt/3dab3535007acf108391649766409421
# usethis::use_git_config(user.name = "vsaulai", user.email = "v.saulai@gmail.com")
# usethis::create_github_token() 
# credentials::set_github_pat("YourPAT")
# gitcreds::gitcreds_set()
```




```{r , echo=FALSE, warning=FALSE}
plot(pressure)
```


