---
title: "R key list of useful commands"
author: "Saulai Vue"
date: "2025-04-03"
output: word_document
---

```{r setup, echo=FALSE, include=FALSE}

```

This list is mainly used to combine all operators, functions, and useful commands in R primarily for practicing in class. Note some straightforward elements I have not put here.

## Accessibility
ls() - shows all current objects in your environment
print() - print out a given command
view() - view a data structure

## R Arithmetic Operators
%%  Modulus (Remainder from Division)
%/%  Integer Division

## Assignment Operators
"my_var <- 3"  "my_var <<- 3"

## R Comparison Operators
==  Equal
!=  Not equal
>   Greater than
<   Less than
>=  Greater than or equal to
<=  Less than or equal to

## Math functions
sqrt()  square root
abs()   absolute value
ceiling()  rounds upwards to nearest integer
floor()    rounds downwards to nearest integer
seq()   ARGUMENTS: seq(from, to, by, length.out) this function generates a sequence of numbers with a specified pattern
            from - starting value of sequence
            to   - ending value of sequence
            by   - step size between numbers
            length.out - total number of elements in sequence

## Checking Data Types
is.vector(your_data)   Checks if this is a vector, returns a boolean(can also check other types, data frames, lists, matrices, factor, arrays, etc.)
class()    Checks the class data type (vector, matrix, list, etc.)
typeof()   Checks the object data type (numeric, character, logical, etc.)

## R Logical Operators
&   Element-wise Logical AND operator. Returns TRUE if both elements are TRUE.
&&  Logical AND operator = Returns TRUE if both statements are TRUE.
|   Element-wise Logical OR operator. Returns TRUE if one of the statements is TRUE.
||  Logical OR operator. Returns TRUE if one of the statements is TRUE.
!   Logical NOT - Returns FALSE if statement is TRUE

## Escape Characters
\   This denotes a character that is followed by another character that is typically illegal that you would want to insert inside a string.
\\	Backslash
\n	New Line
\r	Carriage Return
\t	Tab
\b	Backspace

## R Miscellaneous Operators
:   Creates a series of numbers in a sequence   "x <- 1:10"
%in%  Find out if an element belongs to a vector   "x %in% y" --- Returns a logical vector

%>%  Pipe operator

## List of Package Libraries
```{r , echo=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(tidyverse)
library(ggplot2) #animated visuals
library(ggrepel) #prevent overlapping text labels in ggplot2
library(gganimate) #create animated visuals in ggplot
library(tidyverse)
library(gapminder) #for some data
library(gifski) #creation of gifs, could also use library(png)
library(lubridate) #working with time variables
library(maps) #for mapping with ggplot
library(mapdata) #for mapping with ggplot
library(transformr) #for changing chloropleths across time
library(ggthemes) #extra color options

```

## Conditionals
if (Boolean_conditional) { <body> } -- An "if statement" that only executes if the condition boolean is TRUE.
else if (Boolean_conditional) { <body> } -- A statement "if the previous conditions were not true, then try this condition"

## Loops
while (Boolean_condition) { <body> } -- A loop that executes a set of statements as long as a condition is TRUE.
  break -- statement that can stop a loop even if the condition is TRUE.
  next  -- skips an iteration without terminating the loop
for (variable in sequence) {   - variable: the loop variable that takes on each value in the sequence
  <# Code to execute>}  - sequence: A collection of values (vectors, lists, or ranges) that the loop will iterate over


## Data Structures
Vector - A one-dimensional data structure that is homogeneous(i.e., all elements must be of the same data type)
      temp <- c(1,2,3,4) -- defining a vector starts with the concatenate function.
          indexing- addressing certain elements within a vector or matrix using a bracket [] (note. arguments differ for )
              vector_name[index] - position of the element within a vector that you want to access
              df[["column_name"]] - data frame accessing
              print(df$rowname) - accessing by column extracting
              subset() - accessing a data frame with certain conditions
              

Matrix - A two-dimensional data structure that is also homogeneous (i.e., all elements must be of the same data type)
      matrix(1:6, nrow = 2, ncol = 3) - arguments for making a matrix from a sequence with rows and columns defined.
        colSums(), rowSums() - if specified on a matrix, immediately "squishes" the columns or rows into the dimensions of the table.
        
Array - A multi-dimensional data structure that can be 3D or 4D
      array(), arguments "array(<data>, dim = c(x,y,z)) - where dimensions can be predefined while creating an array
Data Frame - A two-dimensional table-like structure containing multiple types of different data types. Works similar to a spreadsheet.
      data.frame() - can define columns with different data types with this argument
      


## Github

```{r , echo=FALSE, warning=FALSE}
usethis::git_sitrep() # gives information on current github user info
# https://gist.github.com/Z3tt/3dab3535007acf108391649766409421
# usethis::use_git_config(user.name = "vsaulai", user.email = "v.saulai@gmail.com")
# usethis::create_github_token() 
# credentials::set_github_pat("YourPAT")
# gitcreds::gitcreds_set()
```

## Advanced Git

+ branching: you can maintain "parallel" versions of your repository.
This is useful for "exploratory" work. These can later be merged into
the "main" branch.
+ bisecting: you can identify the commit responsible for introducing a
bug using binary search
+ hooks: you can automatically execute tasks based upon git behaviors;
can be used to automate testing/deployment


```{r , echo=FALSE, warning=FALSE}
plot(pressure)
```


