---
title: "R key list of useful commands"
author: "Saulai Vue"
date: "2025-04-03"
output: word_document
---

```{r setup, echo=FALSE, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This list is mainly used to combine all operators, functions, and useful commands in R primarily for practicing in class. Note some straightforward elements I have not put here.

## Accessibility
ls() -  shows all current objects in your environment
rm(object_name) -	 Removes a specific object from the environment.
rm(list = ls()) - 	Clears all objects from the environment.
getwd() -	 Gets the current working directory.
setwd("path/to/directory") - 	Sets the working directory.
dir() or list.files() - 	Lists all files in the current directory.
help(function_name) - Displays help documentation for a function.
?function_name -	Shortcut for the help() function.
example(function_name) -	Shows an example usage of a function.
str(object) -	 Displays the structure of an object.
summary(object) - 	Provides a summary of a dataset, vector, or model.

## R Arithmetic Operators
%%  Modulus (Remainder from Division)
%/%  Integer Division

## Assignment Operators
"my_var <- 3"  "my_var <<- 3"

## R Comparison Operators
==  Equal
!=  Not equal
>   Greater than
<   Less than
>=  Greater than or equal to
<=  Less than or equal to

## Math functions
sqrt()  square root
abs()   absolute value
ceiling()  rounds upwards to nearest integer
floor()    rounds downwards to nearest integer
seq()   ARGUMENTS: seq(from, to, by, length.out) this function generates a sequence of numbers with a specified pattern
            from - starting value of sequence
            to   - ending value of sequence
            by   - step size between numbers
            length.out - total number of elements in sequence

## Checking Data Types
is.vector(your_data) -  Checks if this is a vector, returns a boolean(can also check other types, data frames, lists, matrices, factor, arrays, etc.)
class()  -  Checks the class data type (vector, matrix, list, etc.)
typeof() -  Checks the object data type (numeric, character, logical, etc.)
print(object)	Prints an object’s contents to the console.
View(dataframe)	- Opens a spreadsheet-like view of a data frame.
head(dataframe, n=6) -	Displays the first n rows of a data frame.
tail(dataframe, n=6) -	Displays the last n rows of a data frame.
dim(dataframe) -	Returns the number of rows and columns of a data frame or matrix.
names(dataframe) -	Returns the column names of a data frame.
class(object) -	Returns the data type of an object.
typeof(object) -	Returns the low-level type of an object.
is.na(object) -	Checks for missing values.
complete.cases(dataframe) -	Returns TRUE for rows without missing values.
any(is.na(dataframe)) -	Checks if there are any missing values in the dataset.

## Handling Data
nrow(dataframe) - 	Returns the number of rows in a data frame.
ncol(dataframe)	-  Returns the number of columns in a data frame.
dim(dataframe)	-  Returns both the number of rows and columns.
colnames(dataframe) - 	Returns column names of a data frame.
rownames(dataframe) - 	Returns row names of a data frame.
unique(vector) - 	Extracts unique values from a vector.
sort(vector) - 	Sorts a vector in ascending order.
order(vector) - 	Returns the indices that would sort a vector.
table(vector) - 	Creates a frequency table.

## R Logical Operators
&   Element-wise Logical AND operator. Returns TRUE if both elements are TRUE.
&&  Logical AND operator = Returns TRUE if both statements are TRUE.
|   Element-wise Logical OR operator. Returns TRUE if one of the statements is TRUE.
||  Logical OR operator. Returns TRUE if one of the statements is TRUE.
!   Logical NOT - Returns FALSE if statement is TRUE

## Escape Characters
\   This denotes a character that is followed by another character that is typically illegal that you would want to insert inside a string.
\\	Backslash
\n	New Line
\r	Carriage Return
\t	Tab
\b	Backspace
\
## R Miscellaneous Operators
:   Creates a series of numbers in a sequence   "x <- 1:10"
%in%  Find out if an element belongs to a vector   "x %in% y" --- Returns a logical vector

%>%  Pipe operator (Ctrl + Shift + C) 

## List of Package Libraries
install.packages("package_name") - 	Installs a package.
library(package_name) - 	Loads a package.
update.packages() - 	Updates all installed packages.
sessionInfo() - 	Displays R session details, including loaded packages.
save(object, file = "filename.RData") - 	Saves an R object to a file.
load("filename.RData") - 	Loads an R object from a file.
write.csv(dataframe, "file.csv") - 	Exports a data frame to a CSV file.
read.csv("file.csv") - 	Imports a CSV file into R.

```{r , echo=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(tidyverse)
library(ggplot2) #animated visuals
library(ggrepel) #prevent overlapping text labels in ggplot2
library(gganimate) #create animated visuals in ggplot
library(tidyverse)
library(dslabs)
library(gapminder) #for some data
library(gifski) #creation of gifs, could also use library(png)
library(lubridate) #working with time variables
library(maps) #for mapping with ggplot
library(mapdata) #for mapping with ggplot
library(transformr) #for changing chloropleths across time
library(ggthemes) #extra color options

```

## Conditionals
if (Boolean_conditional) { <body> } -- An "if statement" that only executes if the condition boolean is TRUE.
else if (Boolean_conditional) { <body> } -- A statement "if the previous conditions were not true, then try this condition"

## Loops
while (Boolean_condition) { <body> } -- A loop that executes a set of statements as long as a condition is TRUE.
  break -- statement that can stop a loop even if the condition is TRUE.
  next  -- skips an iteration without terminating the loop
for (variable in sequence) {   - variable: the loop variable that takes on each value in the sequence
  <# Code to execute>}  - sequence: A collection of values (vectors, lists, or ranges) that the loop will iterate over

## Data Structures
Vector - A one-dimensional data structure that is homogeneous(i.e., all elements must be of the same data type)
      temp <- c(1,2,3,4) -- defining a vector starts with the concatenate function.
          indexing- addressing certain elements within a vector or matrix using a bracket [] (note. arguments differ for )
              vector_name[index] - position of the element within a vector that you want to access
              df[["column_name"]] - data frame accessing
              print(df$rowname) - accessing by column or specific name type
              subset() - accessing a data frame with certain conditions
          comparing vectors
             use identical() or all.equal()

Matrix - A two-dimensional data structure that is also homogeneous (i.e., all elements must be of the same data type)
      matrix(1:6, nrow = 2, ncol = 3) - arguments for making a matrix from a sequence with rows and columns defined.
        colSums(), rowSums() - if specified on a matrix, immediately "squishes" the columns or rows into the dimensions of the table.
        
sometimes it is ideal to make any subset of data another object instead of the initial thing.  
        
Array - A multi-dimensional data structure that can be 3D or 4D
      array(), arguments "array(<data>, dim = c(x,y,z)) - where dimensions can be predefined while creating an array
Data Frame - A two-dimensional table-like structure containing multiple types of different data types. Works similar to a spreadsheet.
      data.frame() - can define columns with different data types with this argument

## Data Manipulation

## Escape Sequences (Special Characters)

\n	New line
\t	Tab
\\	Backslash \
\"	Double quote " inside a string
\'	Single quote ' inside a string
\\d  represents digits, inside a string
\s  represents white space, inside a string
^    start of the string
\\s* none or  more white space
$   end of string


## Strings (str) Detect, Locate, Extract, Describe, Replace

str_detect	Detect  	Is the pattern in the string?	grepl
str_which 	Detect  	Returns the index of entries that contain the pattern.	grep
str_subset 	Detect  	Returns the subset of strings that contain the pattern.	grep with value = TRUE
str_locate 	Locate    	Returns positions of first occurrence of pattern in a string.	regexpr
str_locate_all  	Locate  	Returns position of all occurrences of pattern in a string.	gregexpr
str_view   	  Locate  	  Show the first part of the string that matches pattern.	
str_view_all  	Locate  	Show me all the parts of the string that match the pattern.	
str_extract  	Extract 	Extract the first part of the string that matches the pattern.	
str_extract_all 	Extract	  Extract all parts of the string that match the pattern.	
str_match 	  Extract 	  Extract first part of the string that matches the groups and the patterns defined by the groups.	
str_match_all 	Extract 	Extract all parts of the string that matches the groups and the patterns defined by the groups.	
str_sub   	Extract 	Extract a substring.	substring
str_split   	Extract	    Split a string into a list with parts separated by pattern.	strsplit
str_split_fixed	    Extract   	Split a string into a matrix with parts separated by pattern.	strsplit with fixed = TRUE
str_count	    Describe    	Count number of times a pattern appears in a string.	
str_length	    Describe    	Number of character in string.	nchar
str_replace   	Replace   	Replace first part of a string matching a pattern with another.	
str_replace_all	    Replace   	Replace all parts of a string matching a pattern with another.	gsub
str_to_upper	    Replace	    Change all characters to upper case.	toupper
str_to_lower    	Replace   	Change all characters to lower case.	tolower
str_to_title    Replace	    Change first character to upper and rest to lower.	
str_replace_na    	Replace	    Replace all NAs to a new value.	
str_trim    	Replace   	Remove white space from start and end of string.	
str_c   	Manipulate  	Join multiple strings.	paste0
str_conv    	Manipulate	Change the encoding of the string.	
str_sort    	Manipulate	Sort the vector in alphabetical order.	sort
str_order   	Manipulate	Index needed to order the vector in alphabetical order.	order
str_trunc   	Manipulate  	Truncate a string to a fixed size.	
str_pad	    Manipulate  	Add white space to string to make it a fixed size.	
str_dup	    Manipulate  	Repeat a string.	rep then paste
str_wrap  	Manipulate	Wrap things into formatted paragraphs.	
str_interp  	Manipulate  	String interpolation.	sprintf
## Github

```{r, echo=FALSE, warning=FALSE}
# usethis::git_sitrep() # gives information on current github user info
# https://gist.github.com/Z3tt/3dab3535007acf108391649766409421
# usethis::use_git_config(user.name = "vsaulai", user.email = "v.saulai@gmail.com")
# usethis::create_github_token() 
# credentials::set_github_pat("YourPAT")
# gitcreds::gitcreds_set()
```

## Advanced Git

+ branching: you can maintain "parallel" versions of your repository.
This is useful for "exploratory" work. These can later be merged into
the "main" branch.
+ bisecting: you can identify the commit responsible for introducing a
bug using binary search
+ hooks: you can automatically execute tasks based upon git behaviors;
can be used to automate testing/deployment

## basicplots
plot(): generic plotting function
points(): add points to an existing plot
lines(), abline(): add lines to an existing plot
text(), legend(): add text to an existing plot
rect(), polygon(): add shapes to an existing plot
hist(), image(): histogram and heatmap
heat.colors(), topo.colors(), etc: create a color vector
density(): estimate density, which can be plotted
contour(): draw contours, or add to existing plot
curve(): draw a curve, or add to existing plot


## ggplot

